---
title: "Homework # 2"
author: "Orozco Carstensen Ana Paloma"
date: "2/17/2022"
output: html_document
---
<style type = "text / css">
  body {
    font-family: Arial;
  font-size: 10pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem # 1**

Consider the following quantities, discuss and justify what type of probability distribution the following variables can have.

**a.** The size of adult male footwear in Mexico.

    Answer: Normal or unimodal distribution. Because there are very few with very small or very large sizes, almost all of them will be distributed around the average.

**b.** The distribution of inhabitants in human settlements in Mexico.

    Answer: Normal or unimodal distribution. Because hardly anyone lives in isolation; almost all of them are distributed around the average, in groups of neither many nor few houses.

**c.** The average speed of cars in Bernardo Quintana at 09:00am on a weekday.

    Answer: Normal or unimodal distribution. Because it's going to be almost impossible to go too fast or too slow; almost all of them are going to be distributed around the average (which is not going to be very fast, haha).

**d.** The average speed of cars in Bernardo Quintana at 03:00am.

    Answer: Definitely long tail. Most are going to go very fast and very few are going to go slow, they would not be distributed around the average.


## **Problem # 2** 

Let's suppose that we have 2 nodes (A and B) in a network. Node A and B have the same degree, but node A has a clustering coefficient of 1 and node B, low. Draw this situation and explain what would be a better target if you want to eliminate them and affect the network.

First, the draw: 


```{r echo=FALSE}
library(knitr)
library(png)
enfj_path<-"/Users/carstensen/Desktop/Desktop - Ana’s MacBook Pro/Sexto semestre/Functional_genomics/Homeworks/Image.png"
include_graphics(enfj_path)
```


    Discussion: In the case that the network looks like the draw, the better target would be the node connecting both nodes, A and B. In the case that there are more nodes and clusters and both A and B are not connected to the same node (in the draw: not completed purple lines), the better target would be node A, because it would be a hub. 

## **Problem # 3** 

Write a program in R that, using functions, calculates from an adjacency matrix:

First, load a matrix to test the functions: 



```{r}
Friends_Network <- read.csv ("Data/Friends_NW.csv") # Load data 
row.names (Friends_Network) <- Friends_Network [ , 1] # Take the first column and make it the row names
Friends_Network <- Friends_Network [ , -1] # Take first column
as.matrix (Friends_Network) # Make it matrix
```

**a.** Whether it is directed or not.

    If its not directed, the matrix should be symmetric

First, assign the main values 

```{r}
i = 0 # Assign the value of i 
j = 0 # Assign the value of j
Directed_Network <- c () # Create a vector to save the outputs
```

Then, create a loop for 

```{r}
for (i in 1 : nrow (Friends_Network)){ # For i in 1 to the number of rows in the matrix
  for (j in 1 : ncol (Friends_Network)){ # For j in 1 to the number of columns in the matrix
    Directed_Network <- c (Directed_Network, Friends_Network [i, j] != Friends_Network [j, i]) # On the vector, assign TRUE values if its the exact same; FALSE, if not 
  }
}

Directed_Network # Print 
```

Finally, check if all the values are TRUE, which would mean that the matrix is symmetric

```{r}
all (Directed_Network) # Are all the values true ? 
```
    
    False, it is directed


**b.** Whether it is weighted or not.

    If it is weighted, there will be values > 1

```{r}
any (Friends_Network > 1) # Any of the values are > 1 ?
```

    False, it is not weighted



**c.** The degree of the nodes.

First, assign the main values 

```{r}
i = 0 # Assign the value of i 
In_Degrees_FN <- c () # Create a vector to save the outputs
```

Then, create a loop for 

```{r}
for (i in 1 : nrow (Friends_Network)){ # For i in 1 to the number of rows in the matrix
  In_Degrees_FN <- c (In_Degrees_FN, sum (Friends_Network [i , ])) # On the vector, save the sums of the rows, one by one
}

In_Degrees_FN # Print 
``` 


**d.** The histogram of the degree.

``` {r}
hist (In_Degrees_FN, main = "Histogram of degree distribution", xlab = "In degrees") # Create the histogram
``` 


## **Problem # 4** 

From the protein interaction network called “yeast”, write a program in R that:

**a.** Calculate the distribution of connectivities and plot that distribution.

First, load the data:

```{r}
library (igraph) # Load library 

library (igraphdata) # Load library 

data ("yeast") # Load data

```

Then, calculate the degrees and degree distribution:

```{r}
Yeast_Degree <- degree (yeast) # Check the degrees
head (Yeast_Degree) # Only look at the first ones

Yeast_D_D <- degree_distribution (yeast) # Check the degree distribution
Yeast_D_D # Print
```

Finally, the plot: 

``` {r}
plot (degree_distribution (yeast), main = "Degree distribution", xlab = "Degree", ylab = "Frequency") # Plot
```
   
    Discussion: It is definetely a long tail network, a lot of the nodes have a few connections, but a few nodes have a lot connections. 


**b.** Plot the boxplot of the connectivity distribution and discuss the result.

```{r}
Boxplot_DD_Y <- boxplot (Yeast_D_D, main = "Degree distribution") # Make the boxplot
```

    Discussion: It is long tail because almost all the nodes have a few connections (median inside de box) and just a few, a lot. The distribution is not normal. 

**c.** Find the proportion of nodes that have 15 or more connectivities.

First, find the length of all the data:

```{r}
Length_D <- length (Yeast_Degree) # Find the length of the complete data
Length_D # Print
```

Then, find which nodes have 15 or more connections and the total of those nodes:

```{r}
More_15 <- which (Yeast_Degree >= 15) # Find which nodes have 15 or more connections (the numbers are the position, not the connections)
More_15 # Print

Length_15 <- length (More_15) # Find the length of the nodes that have 15 or more connections
Length_15 # Print
```

Finally, calculate the proportion: 

```{r}
Proportion_15 <- (Length_15 * 100) / Length_D # Calculate the proportion
Proportion_15 # Print
```


**d.** Calculate the maximum degree of the network.

```{r}
Max_Degree <- max (Yeast_Degree) # Obtain the maximum degree
Max_Degree # Print
```

**e.** Calculate the diameter.

```{r}
Diameter_Y <- diameter (yeast) # Calculate the diameter 
Diameter_Y # Print
```

**f.** The longest path.

```{r}
Longest_Path_Y <- get_diameter(yeast) # Obtain the longest path (15)
Longest_Path_Y # Print
```

    The longest path is: YPL259C YPR029C YLR170C YGL206C YER177W YDR028C YGL115W YDR422C YHR114W YBL007C YNL271C YLR190W YIL050W YPL031C YER059W YDL224C, with 16 proteins, 15 steps


**g.** Remove the 10 most connected nodes of the network and determine the diameter each time you do so.

First, assign the main values: 

```{r}
i = 0 # Assign the value of i 
First_10 <- sort (Yeast_Degree, decreasing = TRUE) [1:10]  # Sort and select the first 10 nodes (the ones with more connections)
First_10 # Print
Diameter_Yeast <- c () # Create a vector to save the diameters 
```

Then, create a loop for: 

```{r}
for (i in 1 : 10) { # For i in 1 to 10
  Diameter_Yeast [i] <- diameter (yeast) # Calculate the diameter for network
  yeast <- delete_vertices (yeast, sample (First_10 [i])) # Delete the node corresponding to i 
}
```

Finally, obtain the diameters on the vector: 

```{r}
print (Diameter_Yeast) # Print
```

Reload the data, for further problems 

```{r}
data ("yeast") # Load data
```

**h.** Determine the ten most important nodes by at least 3 measures of centrality.

1. Degree centrality 

```{r}
Y_Degree_centrality <- degree (yeast) # Calculate the degree of the network
sort (Y_Degree_centrality, decreasing = TRUE) [1:10] # Order the data
```

    Discussion: Those are the 10 most important proteins. It is because those proteins are the ones with more connections. 


2. Eigenvector centrality

```{r}
Y_Eigenvector_Centrality <- eigen_centrality (yeast) # Calculate the eigenvector centrality of the network
sort (Y_Eigenvector_Centrality $ vector, decreasing = TRUE) [1:10] # Order the data
```

    Discussion: Those are the 10 most important proteins. It is because those proteins are the ones with more influence on the network


3. Closness centrality 

```{r}
Y_Closness_centrality <- closeness (yeast) # Calculate the closness centrality of the network
sort (Y_Closness_centrality, decreasing = TRUE) [1:10] # Order the data
``` 
  
    Discussion: Those are the 10 most important proteins. It is because those proteins are the closest to other nodes


**i.** Cluster the network by at least two methods and determine the size of the largest cluster.

1. Cluster by community structure detecting based on the leading eigenvector of the community matrix

First, create the clusters:

```{r}
Cluster_LE_Y <- cluster_leading_eigen (yeast) # Create the cluster
Cluster_LE_Y # Print
```

Then, obtain the information:

```{r}
membership (Cluster_LE_Y) # Obtain which member are in which clusters
table (membership (Cluster_LE_Y)) # Obtain the number of members by clusters

sort (table (membership (Cluster_LE_Y)), decreasing = TRUE) # Sort the members of the cluster 
```

    The larger cluster is cluster number 96, which has 1,483 proteins


2. Cluster via greedy optimization of modularity 

First, create the clusters:

```{r}
Cluster_FG_Y <- cluster_fast_greedy (yeast) # Create the cluster
Cluster_FG_Y# Print
```

Then, obtain the information:

```{r}
membership (Cluster_FG_Y) # Obtain which member are in which clusters
table (membership (Cluster_FG_Y)) # Obtain the number of members by clusters

sort (table (membership (Cluster_FG_Y)), decreasing = TRUE) # Sort the members of the cluster 
```

    The larger cluster is cluster number 1, which has 744 proteins

Finally, compare the clusters:

```{r}
Comparation <- compare (Cluster_LE_Y, Cluster_FG_Y, method = "adjusted.rand") # Compare clusters
Comparation # Print
```

    0.1601622 = not similar


**j.** Determine if the network is small world, ultra small world, or other.

Create an histogram to see the distribution of the nodes and their conections 

```{r}
hist (Yeast_D_D) # Create histogram
```
    
    Discussion: it is definetely long-tail: almost all of the nodes have 0.0 - 0.0 5% conections, while only over 1-3, have 0.25 - 0.30 % conections. If it were small-world, there would be a high local clustering, like all nodes connected to all nodes. 



